Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-06-23 ltx/PLoS/LaTeX_Manuscript.tex
added pdf plots in the required format

The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 
\begin{equation}
	\mathbf W_0^{(m)}=\frac{\lambda}{ L+\lambda} 
\end{equation}
\begin{equation}
	\mathbf W_0^{(c)}=\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta) 
\end{equation}
\begin{equation}
	\mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L, 
\end{equation}
where the superscripts $m$ and $c$ stand for mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the cross-covariance matrix of the states, $\mathbf M$, is also computed. A summary of the steps in the URTSS algorithm is given in Table~\ref{tab:UKFAlgorithm}. It should be noted that the disturbance, $\mathbf{e}_t$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used, instead of augmenting the state vector with the noise terms with the added benefit of reducing the dimension and the number of sigma points required in the algorithm.

@@

The sensors were spaced on a regular square $14 \times 14$ grid, the sensor width was set to $\sigma^2_m = 0.81$. This provides a width of $1.5$~mm at half the maximum amplitude of the basis function, with spacing $\Delta_y = 1.5$~mm. This arrangement modelled the recording as having some cross talk, which is typical for neurophysiological recordings.
     		\includegraphics[scale=1]{./Graph/ExperimentalFigure.pdf}
    	\caption{\textbf{a.} Example of the neural feld. \textbf{b.} Example of data generated by the full neural field model from the first 5 sensors. \textbf{c.} Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 
 \label{fig:experimental design}
\subsection*{Spatial Frequency Analysis} 
Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT} shows the spatial frequency of the neural field. The cut-off frequency of the neural field is $\nu_c \approx 0.24$~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of $2.05$~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to prevent problems associated with spatial aliasing.
% \begin{figure}
% \centering
% \subfigure[][]{\label{fig:FieldFFT}\includegraphics{./Graph/FFTTrueField.pdf}}
% \subfigure[][]{\label{fig:EstimatedFieldFFT}\includegraphics{./Graph/FFTEstimatedField.pdf}}\\
% \caption{Spatial frequency analysis of the neural field. \textbf{a.} The average (over time) power in dB of the spatial frequency of the neural field. \textbf{b.} The average (over time) power in dB of the spatial frequency of the reconstructed neural field from the basis function decomposition. The reconstructed field shows a narrower bandwidth as a result of the band limiting properties of the sensors and basis functions.}
% \label{fig:FFTTrueEstimate}
% \end{figure}
The spatial frequency of the observed field is shown in Figure~\ref{fig:ObservationFreqAnal}. The cut-off frequency is $\nu_{cy} \approx 0.2$~Hz ($-3$~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:SensorFunctionFreqResponse}. Using an oversampling parameter of $\rho_{\phi}=1$, the minimum distance between the field basis functions was calculated to be $2.5$~mm to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). The distance between basis functions was set to $\Delta_{\phi}=2.5$~mm, which is sufficient to capture the significant dynamics according to our criteria.
Given the basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$~mm. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\mathbf{\nu_{c\phi}} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:BasisFunctionFreqResponse}.
% \begin{figure} 
% 	\begin{center}
% 		\subfigure[][]{\label{fig:ObservationFreqAnal}\includegraphics{./Graph/ObservationFrequencyResponse.pdf}}\\	
% 		\subfigure[][]{\label{fig:SensorFunctionFreqResponse}\includegraphics{./Graph/SensorChar.pdf}}
% 		\subfigure[][]{\label{fig:BasisFunctionFreqResponse}\includegraphics{./Graph/BasisChar.pdf}}
% 	\end{center}
% 	\caption{Spatial frequency analysis. The dashed line shows the cut-off frequency (-3 dB point). \textbf{a.} Two-dimensional spatial frequency in power dB, averaged over 400 time steps) of the observed field. The symmetry of the sensors and field basis functions allow for the one-dimensional representation of the spectral characteristics. \textbf{b.} The spatial frequency response of a one-dimensional model sensor. \textbf{c.} The spatial frequency response of a one-dimensional neural field basis function, illustrating a narrower band width than the sensors.} 
% \end{figure}

@@

Histograms of the parameter estimates obtained over the 100 realisations are shown in Figure~\ref{fig:Parameters}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good concordance with the actual values. 
% It should be noted that when the excitatory connectivity parameters, $\hat\theta_0$ and $\hat\theta_2$, were under (or over) estimated, the inhibitory connectivity parameter, $\hat\theta_1$, tended to be over (or under) estimated. This is demonstrated in Figure~\ref{fig:ParametersRatio} where we show the distribution of the ratios of the parameter estimates. 
\begin{table}
\centering
\begin{tabular}{cccc}
	\hline\hline Parameter & Actual Value & Mean Estimate $\pm$ Std. Dev. & Bias (\%) \\
	\hline\hline
 	$\xi$ & $0.9$ & $0.925\pm 0.004$ & $2.778$ \\
	$\theta_0$ & $10$ & $10.228\pm2.794$ & $2.28$\\
	$\theta_1$ & $-8$ & $-8.495\pm1.913$ & $-6.187$\\
	$\theta_2$ & $0.5$ & $0.545\pm0.077$ & $9$\\
	% $\mid\theta_0/\theta_1\mid$ &$1.25 $ & $1.189\pm 0.0812$ &$4.88$\\
	% $\theta_0/\theta_2$ &$20$ & $18.567\pm3.757$ &$7.165$ \\
\hline\hline
\end{tabular}
\caption{Actual parameter values compared to parameter estimates.}
\label{tab:Parameters estimates}
\end{table}
% \begin{figure}
%     \centering
%     \subfigure[][]{\includegraphics{./Graph/zeta.pdf}}
%     \subfigure[][]{\includegraphics{./Graph/theta0.pdf}}\\
%     \subfigure[][]{\label{fig:theta1dis}\includegraphics{./Graph/theta1.pdf}}
%     \subfigure[][]{\includegraphics{./Graph/theta2.pdf}}
% \caption{Distribution of the parameter estimates over 100
% realisations. The solid line shows the actual value of the parameters and the dotted line shows the mean of the estimated parameters. \textbf{a.} The distribution of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant. \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$.}
% \label{fig:Parameters}
% \end{figure}
% \begin{figure}
%     \centering
% \subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
% \subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
% \caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20). In each case, the solid and dotted lines show the locations of the true and estimated ratios respectively.}
% \label{fig:ParametersRatio}
% \end{figure}
The reconstructed connectivity kernels from the estimated parameters are illustrated in Figure~\ref{fig:KernelEstimates}, where the mean over the 100 realisations is shown as a dotted line and the $95\%$ confidence interval is indicated by the grey shaded region. The figure verifies that the shape of the estimated connectivity kernel is in good accordance to the true kernel. The true kernel lies within the confidence interval; however, the mean estimated kernel is more negative than the actual kernel due to the small bias on the estimates of the inhibitory parameter, $\hat\theta_1$, as seen in Figure~\ref{fig:theta1dis}.

@@

% \begin{figure}
% \centering 
% \subfigure[][]{\label{fig:TrueField}\includegraphics{./Graph/TrueField90.pdf}}
% \subfigure[][]{\includegraphics{./Graph/EstimatedField90.pdf}}
% \subfigure[][]{\includegraphics{./Graph/EstimationError90.pdf}}\\
% \caption{Examples of the actual and estimated neural fields and the error. \textbf{a.} The actual field at a single time instant from the full neural field model that was used to generate data. \textbf{b.} The reconstructed field of the reduced model, showing the effect of the basis function decomposition on the frequency content of the estimated field. \textbf{c.} The absolute error between the true and reconstructed fields. The error shows high frequency behaviour.}
% \label{fig:FieldEstimate}
% \end{figure}

@@

\begin{table}[!ht]
\caption{
\bf{Notation and Parameters}}
\begin{tabular}{|c|c|c|c|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Value} & \textbf{Units}\\
	\hline
\multicolumn{4}{|c|}{\emph{Domain and indices}}\\
\hline
$\Omega$ & spatial domain & &\\
$t$ & time & & seconds\\
$\mathbf{r}$ & spatial location & &\\
$n$ & sensor index $n=1,..,N$ & &\\
$T_s$ & time step & $0.001$ & seconds\\
$T$ & number of time steps & 0.4 & seconds\\
\hline
\multicolumn{4}{|c|}{\emph{Spatiotemporal signals}}\\
\hline
    $y(\mathbf{r}_n,t)$ & observation & & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & & mV \\
	$g(\mathbf{r},t)$ & weighted firing rate function & & \\
	$f(\mathbf{r},t)$ & firing rate function & & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, (covariance $\gamma$) & &\\
	$\varepsilon(\mathbf{r}_n,t)$ & observation noise, (covariance $\Sigma_\epsilon$) & &\\
\end{tabular}
\begin{flushleft}Table caption
\end{flushleft}
\label{tab:label}
\end{table}

% ~~~~~~~~~~~~~~~~~~~~

% 
% \begin{table}[h!]\footnotesize
%     \begin{tabular}{ll}
%         	$h(t)$ & post-synaptic response kernel \\
% 	$m(\mathbf{r},\mathbf{r}')$ & sensor kernel, width $\sigma_m^2$ \\
% 	$\eta(t)$ & Heaviside function \\
% 	$\zeta$ & inverse synaptic time constant \\
% 	$w(\mathbf{r},\mathbf{r}')$ & spatial connectivity kernel \\
% 	$f_{max}$ & maximal firing rate \\
% 	$\varsigma$ & slope of sigmoidal activation function \\
% 	$v_0$ & firing threshold \\
% 	$\delta(t)$ & Dirac-delta function \\
% 	$D$ & Temporal differential operator \\
% 	$\xi$ & time constant parameter
%     \end{tabular}
%     \caption{Model}
% \end{table}
% 
% \begin{table}[h!]\footnotesize
%     \begin{tabular}{ll}
%         $L$ & number of field basis functions  \\
%     	$\mathbf{\phi(r)}$ & vector of Gaussian basis functions \\
%     	$\mathbf{x}_t$ & state vector at time $t$ \\
%     	$\mathbf{\psi}$ & vector of connectivity kernel basis functions \\
%     	$\theta$ & vector of connectivity kernel parameters \\
%     	$\Gamma$ & inner product of field basis functions \\
%     	$q()$ & state function \\
%     	$Q()$ & state function \\
% 	$\gamma(\mathbf r-\mathbf r')$& field disturbance covariance function\\
%     	$\mathbf{e}_t$ & state disturbance \\
% 	$\Sigma_e$ &state disturbance covariance matrix\\
% 	$\mathbf{\epsilon}_t$ & observation noise \\
% 	$\Sigma_{\epsilon}$ &observation noise covariance matrix\\
%     	$\mathbf{C}$ & observation matrix \\
% 	$c_{ij}$& elements of observation matrix
%     \end{tabular}
%     \caption{Decomposition}
% \end{table}
% 
% \begin{table}[h!]\footnotesize
%     \begin{tabular}{rl}
%         $V(\nu)$ & Spectrum of the dynamic field \\
% 	$\nu$ & Spatial frequency \\
% 	$\nu_c$ & Spatial cut-off frequency \\
% 	$\Delta_y$ & Distance between adjacent sensors \\
% 	$\rho$ & Over-sampling parameter \\
% 	$\Delta_{\phi}$ & Distance between field basis functions \\
% 	$\sigma_{\nu}^2$ & Variance of FT of Gaussian basis function \\
% 	$\sigma^2_{\phi}$ & Spatial variance of Gaussian in spatial domain \\
%     \end{tabular}
%     \caption{Frequency Analysis}
% \end{table}
% 
% \begin{table}[h!]\footnotesize
%     \begin{tabular}{ll}
%         $\hat{\mathbf{x}}$ & State estimate \\
% 	$\chi$ & Matrix of sigma vectors \\
% 	$\lambda$& Scaling parameter\\
% 	$\alpha$& Determines the spread of sigma points\\
% 	$\beta$& Prior knowledge of the sigma points distribution\\
% 	$\kappa$& Scaling parameter\\
% 	$W$& Sigma vecors weights\\
%     	$\hat{\mathbf{x}}_t^{f-}$ & Forward prior estimate \\
%     	$\hat{\mathbf{x}}_t^f$ & Forward posterior estimate \\
%     	$\hat{\mathbf{x}}_t^{b-}$ & Backward prior estimate \\
%     	$\hat{\mathbf{x}}_t^{b}$ & Backward posterior estimate \\
%     	$P^f_t$ & Forward posterior covariance matrix \\
%     	$P^{f-}_t$ & Forward prior covariance matrix \\
%     	$P^b_t$ & Backward posterior covariance matrix \\
%     	$P^{b-}_t$ & Backward prior covariance matrix \\
% 	$\mathcal K_{t} $& Filter gain\\
% 	$M_t$& Cross-covariance matrix\\
% 	$S_t$& Smoother gain
%     \end{tabular}
%     \caption{State Estimation}
% \end{table}

>>>>

The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 
	\mathbf W_0^{(m)}=\frac{\lambda}{ L+\lambda} 
	\mathbf W_0^{(c)}=\frac{\lambda}{ L+\lambda}+(1-\alpha^2+\beta) 
	\mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L, 
where the superscripts $m$ and $c$ stand for mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the cross-covariance matrix of the states, $\mathbf M$, is also computed. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that the disturbance, $\mathbf{e}_t$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used, instead of augmenting the state vector with the noise terms with the added benefit of reducing the dimension and the number of sigma points required in the algorithm.
\begin{algorithm}
	\begin{small}
	\caption{The Unscented RTS Smoother}\label{UKFAlgorithm} 
	\begin{algorithmic}[1] 
		\State Forward initialisation 
		\begin{equation*}
		 \hat{\mathbf x}_0, \mathbf P_0 
		\end{equation*}
		\State Forward iteration: for $t \in \left\lbrace 0,\cdots, T\right\rbrace $,
		calculate the sigma points $\mathcal X_{i,t}^f$ using equations \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propagate through equation~\ref{eq:QmatrixForSigmapoints}
% 		\begin{small}
		\begin{equation*}
			\mathcal X_{i,t+1}^{f-}=Q(\mathcal X_{i,t}^f) \quad i=0, \dots, 2L
		\end{equation*}
% 		\end{small}
		calculate the predicted state and the predicted covariance matrix
		\begin{equation*}
			\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{f-} 
		\end{equation*}
		\begin{equation*}
			\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e 
		\end{equation*}
		compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman Filter update equations
		\begin{equation*}
			\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_{\epsilon})^{-1} 
		\end{equation*}
		\begin{equation*}
			\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-}) 
		\end{equation*}
		\begin{equation*}
			\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-} 
		\end{equation*}
		\State Backward initialisation 
		\begin{equation*}
			\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T 
		\end{equation*}
		\State Backward iteration: for $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $ calculate the sigma points $\mathcal X_{i,t}^b$ and propagate them through equation \ref{eq:QmatrixForSigmapoints}
		\begin{equation*}
			\mathcal X_{i,t+1}^{b-}=Q(\mathcal X_{i,t}^b) \quad i=0, \dots, 2L
		\end{equation*}
		 compute the predicted state, the predicted covariance matrix and the cross-covariance matrix
		\begin{equation*}
			\hat{\mathbf x}_{t+1}^{b-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{b-} 
		\end{equation*}
		\begin{equation*}
			\mathbf P_{t +1}^{b-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})^\top+\boldsymbol \Sigma_e 
		\end{equation*}
		\begin{equation*}
			\mathbf M_{t +1}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t}^{b-}-\hat{\mathbf x}_{t}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t+1}^{b-})^\top 
		\end{equation*}
		 Compute the smoother gain, the smoothed state and the smoothed covariance matrix
		\begin{equation*}
			\mathbf S_t=\mathbf M_{t +1}\left[ \mathbf P_{t +1}^{b-}\right] ^{-1} 
		\end{equation*}
		\begin{equation*}
			\hat{\mathbf x}_t^b=\hat{\mathbf x}_t^f+\mathbf S_t\left[\hat{\mathbf x}_{t+1}^{b}-\hat{\mathbf x}_{t+1}^{b-}\right] 
		\end{equation*}
		\begin{equation*}
			\mathbf P_{t}^{b}=\mathbf P_{t}^{f}+\mathbf S_t\left[\mathbf P_{t+1}^{b}-\mathbf P_{t+1}^{b-} \right]\mathbf S_t^\top 
		\end{equation*}
	\end{algorithmic}
\end{small}
\end{algorithm}

@@

The sensors were spaced on a regular square $14 \times 14$ grid, the sensor width was set to $\sigma^2_m = 0.81$. This provides a width of $1.5$~mm at half the maximum amplitude of the basis function, with spacing $\Delta_y = 1.5$~mm. This arrangement modelled the recording as having some cross talk, which is typical for neurophysiological recordings.
 \begin{figure}
    	\begin{center}
     		\includegraphics{./Graph/ExperimentalFigure.pdf}
    	\end{center}
    	\caption{\textbf{a.} Example of the neural feld. \textbf{b.} Example of data generated by the full neural field model from the first 5 sensors. \textbf{c.} Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 
 \label{fig:experimental design}
 \end{figure}
\subsection*{Spatial Frequency Analysis} 
Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT} shows the spatial frequency of the neural field. The cut-off frequency of the neural field is $\nu_c \approx 0.24$~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of $2.05$~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to prevent problems associated with spatial aliasing.
\begin{figure}
\centering
\label{fig:FieldFFT}\includegraphics{./Graph/FFTField.pdf}
\caption{Spatial frequency analysis of the neural field. \textbf{a.} The average (over time) power in dB of the spatial frequency of the neural field. \textbf{b.} The average (over time) power in dB of the spatial frequency of the reconstructed neural field from the basis function decomposition. The reconstructed field shows a narrower bandwidth as a result of the band limiting properties of the sensors and basis functions.}
\label{fig:FFTTrueEstimate}
\end{figure}
The spatial frequency of the observed field is shown in Figure~\ref{fig:ObservationFreqAnal}. The cut-off frequency is $\nu_{cy} \approx 0.2$~Hz ($-3$~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:SensorFunctionFreqResponse}. Using an oversampling parameter of $\rho_{\phi}=1$, the minimum distance between the field basis functions was calculated to be $2.5$~mm to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). The distance between basis functions was set to $\Delta_{\phi}=2.5$~mm, which is sufficient to capture the significant dynamics according to our criteria.
Given the basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$~mm. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\mathbf{\nu_{c\phi}} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:BasisFunctionFreqResponse}.
\begin{figure} 
	\begin{center}
		\label{fig:ObservationFreqAnal}\includegraphics{./Graph/SensorBasisObsFrequencyResponse.pdf}	
	\end{center}
	\caption{Spatial frequency analysis. The dashed line shows the cut-off frequency (-3 dB point). \textbf{a.} Two-dimensional spatial frequency in power dB, averaged over 400 time steps) of the observed field. The symmetry of the sensors and field basis functions allow for the one-dimensional representation of the spectral characteristics. \textbf{b.} The spatial frequency response of a one-dimensional model sensor. \textbf{c.} The spatial frequency response of a one-dimensional neural field basis function, illustrating a narrower band width than the sensors.} 
\end{figure}

@@

Histograms of the parameter estimates obtained over the 100 realisations are shown in Figure~\ref{fig:Parameters}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good concordance with the actual values. 
% It should be noted that when the excitatory connectivity parameters, $\hat\theta_0$ and $\hat\theta_2$, were under (or over) estimated, the inhibitory connectivity parameter, $\hat\theta_1$, tended to be over (or under) estimated. This is demonstrated in Figure~\ref{fig:ParametersRatio} where we show the distribution of the ratios of the parameter estimates. 
\begin{table}
\centering
\begin{tabular}{cccc}
	\hline\hline Parameter & Actual Value & Mean Estimate $\pm$ Std. Dev. & Bias (\%) \\
	\hline\hline
 	$\xi$ & $0.9$ & $0.925\pm 0.004$ & $2.778$ \\
	$\theta_0$ & $10$ & $10.228\pm2.794$ & $2.28$\\
	$\theta_1$ & $-8$ & $-8.495\pm1.913$ & $-6.187$\\
	$\theta_2$ & $0.5$ & $0.545\pm0.077$ & $9$\\
	% $\mid\theta_0/\theta_1\mid$ &$1.25 $ & $1.189\pm 0.0812$ &$4.88$\\
	% $\theta_0/\theta_2$ &$20$ & $18.567\pm3.757$ &$7.165$ \\
\hline\hline
\end{tabular}
\caption{Actual parameter values compared to parameter estimates.}
\label{tab:Parameters estimates}
\end{table}
\begin{figure}
    \centering
    \includegraphics{./Graph/ParametesDistributions.pdf}
  \caption{Distribution of the parameter estimates over 100
realisations. The solid line shows the actual value of the parameters and the dotted line shows the mean of the estimated parameters. \textbf{a.} The distribution of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant. \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$.}
\label{fig:Parameters}
\end{figure}
% \begin{figure}
%     \centering
% \subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
% \subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
% \caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20). In each case, the solid and dotted lines show the locations of the true and estimated ratios respectively.}
% \label{fig:ParametersRatio}
% \end{figure}
The reconstructed connectivity kernels from the estimated parameters are illustrated in Figure~\ref{fig:KernelEstimates}, where the mean over the 100 realisations is shown as a dotted line and the $95\%$ confidence interval is indicated by the grey shaded region. The figure verifies that the shape of the estimated connectivity kernel is in good accordance to the true kernel. The true kernel lies within the confidence interval; however, the mean estimated kernel is more negative than the actual kernel due to the small bias on the estimates of the inhibitory parameter, $\hat\theta_1$, as seen in Figure~\ref{fig:theta1dis}.

@@

\begin{figure}
\centering 
\label{fig:TrueField}\includegraphics{./Graph/FieldTrueEstimationError.pdf}
\caption{Examples of the actual and estimated neural fields and the error. \textbf{a.} The actual field at a single time instant from the full neural field model that was used to generate data. \textbf{b.} The reconstructed field of the reduced model, showing the effect of the basis function decomposition on the frequency content of the estimated field. \textbf{c.} The absolute error between the true and reconstructed fields. The error shows high frequency behaviour.}
\label{fig:FieldEstimate}
\end{figure}

@@

\begin{table}[!ht]
\caption{
\bf{The Unscented RTS Smoother}}
\begin{tabular}{|c|}\hline
\multicolumn{1}{|p{16cm}|}{ 1. Forward initialisation} \\ 
\\
$\hat{\mathbf x}_0, \mathbf P_0$ \\
\\
\multicolumn{1}{|p{16cm}|}{   Forward iteration: for $t \in \left\lbrace 0,\cdots, T\right\rbrace $, calculate the sigma points $\mathcal X_{i,t}^f$ using equations \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propagate through equation~\ref{eq:QmatrixForSigmapoints}} \\
\\
	$\mathcal X_{i,t+1}^{f-}=Q(\mathcal X_{i,t}^f) \quad i=0, \dots, 2L$\\
	\\
\multicolumn{1}{|p{16cm}|}{   Calculate the predicted state and the predicted covariance matrix} \\
\\
$\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{f-} $ \\
\\
$\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e$ \\ 
\\
\multicolumn{1}{|p{16cm}|}{Compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman filter update equations} \\
\\
$\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_{\epsilon})^{-1}$\\
\\
$\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-}) $\\
\\
$\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}$\\ 
\\\hline
\multicolumn{1}{|p{16cm}|}{2. Backward initialisation}\\
\\
$\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T$ \\
\\

\end{tabular}
\begin{flushleft}Table caption
\end{flushleft}
\label{tab:label}
\end{table}
